---
title: "Preparing and exploring Data"
---

## Catch up

```{r}
library(dplyr)
library(dbplyr)
library(DBI)

con <- dbConnect(
  odbc::databricks(),
  HTTPPath = "/sql/1.0/warehouses/300bd24ba12adf8e"
)

```

## 01 - Selecting variables
*Simple strategies to order, and reduce, data to work with*

1. Load the `customer` table to a variable called `customer` 

```{r}

```

2. Select all columns that end with "key"

```{r}

```

3. Move all columns that end with "key" to the front

```{r}

```

4. Select all columns that **do not** end with "key"

```{r}

```


## 02 - Join to tables
**Using left_join() to relate two tables**

1. Load the `nation` table into a variable called the same

```{r}

```

2. Use `left_join` to relate `customer` with `nation` using the nation key 

```{r}

```

3. What are the 5 countries with the most customers?

```{r}

```

## 03 - Prepare base 
*Building the base variable/query*

1. Load the `orders` table in a variable called `orders`

```{r}

```

2. Join `orders` to the `customer` variable (table). Relate them on the
`o_custkey` and `c_custkey` fields.

```{r}

```

3. Join the `nation` variable/table to the `orders` and `customer` variables. 
Use the `c_nationkey` and the `n_nationkey` to relate them.

```{r}

```

4. Load the resulting code into a variable called `rel_orders`. **We do this 
so to get autocomplete working**

```{r}

```

5. Create new columns for the year of the order date, and another for the
month of the order date. Name them `order_year` and `order_month` respectively.

```{r}

```

6. Remove any columns that end in "comment", and end in "key"

```{r}

```

7. Rename `o_custkey` to `customer`, insert code before the selection

```{r}

```

8. Load resulting code to a variable called `prep_orders`

```{r}

```

9. Preview `prep_orders` using `glimpse()`

```{r}

```


## 04 - Answering questions
*Using the base query to answer more complex questions*

1. What are the top 5 countries for total amount ordered?

```{r}

```

2. What are the top 5 countries for total amount ordered for 1998?

```{r}

```

3. What has been the top (1) country, in orders, by year?

```{r}

```

4. Who are the top 5 customers by amount ordered?

```{r}

```

5. What is the country, and market segment, of the top 5 customers by amount ordered?

```{r}

```
