---
title: "Visualizations"
---

## Catch up 

```{r}
library(dplyr)
library(dbplyr)
library(DBI)

con <- dbConnect(
  odbc::databricks(),
  HTTPPath = "/sql/1.0/warehouses/300bd24ba12adf8e"
)

orders <- tbl(con, I("workshops.tpch.orders"))
customers <- tbl(con, I("workshops.tpch.customer"))
nation <- tbl(con, I("workshops.tpch.nation"))

prep_orders <- orders |> 
  left_join(customers, by = c("o_custkey" = "c_custkey")) |> 
  left_join(nation, by = c("c_nationkey" = "n_nationkey")) |> 
  mutate(
    order_year = year(o_orderdate), 
    order_month = month(o_orderdate)
  ) |> 
  rename(customer = o_custkey) |> 
  select(-ends_with("comment"),  -ends_with("key"))
```

## 01 - Auto-collect
**See how `ggplot2` auto-collects data before plotting**

1. Load `ggplot2`

```{r}

```

2. Plot the `n_name` over `n_region_key` from the `nation` table. Use the
column geom.

```{r}

```

## 02 - Plot data

1. Using `prep_order`, pull the total sales by year (`o_totalprice`)

```{r}

```

2. Add to the code, a step to plot the data. Use a column geom

```{r}

```

3. Download the results to R to a variable called `sales_by_year`

```{r}

```

4. Preview `sales_by_year`

```{r}

```

5. Use `sales_by_year` to create the same plot

```{r}

```

6. An example of what multiple iterations of the plot would result in

```{r}

```

## 03 - Plot data by country

1. Create a variable called `country`, with the value "FRANCE"

```{r}

```


2. Modify `sales_by_year`, by adding a `filter` step to have the `n_name` match
the value of `country`

```{r}

```

3. Copy and use the same code from the finalized plot. Add a subtitle with the
value of `country`

```{r}

```

## 04 - Plot data by month

1. Create a new variable called `year`, load it with the value of 1998

```{r}

```

2. Using the same structure, create a new variable called `sales_by_month`. In
addition to country, the filter should include the `order_year`. Group by 
`order_month`

```{r}

```

3. Create the same finalized plot, but using `sales_by_month`. Make sure to
update the the axis, and aesthetics. Also, expand the subtitle to include
the `year` as well.

```{r}

```




