---
title: "Working with LLMs in Databricks"
---

## Catch up

```{r}
library(dplyr)
library(dbplyr)
library(DBI)

con <- dbConnect(
  odbc::databricks(),
  HTTPPath = "/sql/1.0/warehouses/300bd24ba12adf8e"
)
```

## 01 - Accessing AI functions
*Use the sentiment classification function*

1. Create a quick review table using the following code:

```{r}
#| eval: true

reviews <- tribble(
  ~name,     ~review, 
  "adam",    "This is the best toaster I have ever bought",
  "berry",   "Toaster arrived broken, waiting for replancement",
  "charles", "The washing machine is as advertised, can't wait to use it",
  "dan",     "Not sure how to feel about this tevelision, nice brightness but bad definition"
) |> 
  select(review) 
```

2. Copy the `reviews` data frame to your SQL session. Assign it to a variable
called `tbl_reviews`

```{r}
```

3. Create a new field called "sentiment", use `ai_analyze_sentiment()` to 
analyze the "review" field

```{r}

```

## 02 - Specify array
*Using array() to run the classification function*

1. Use `ai_classify()` to find out if we need to follow up with customer. The 
two options should be: 'order complete', and 'need follow up'. Use `array()`
as if you would be using the `c()` function. Name the new field "follow_up"

```{r}

```

2. Add a step that keeps only those orders that need follow up

```{r}

```

## 03 - Process complex output
*Working STRUCT output from an 'ai' function*

1. Use `ai_extract()` to pull the type of product being referred to in the
review. Pass 'product' as the extract argument, and pass it inside an `array()`
call. Name the new field "product"

```{r}

```


2. Append a `compute()` step, and assign to a new variable called `tbl_review`

```{r}

```

3. Preview `tbl_review`

```{r}
 
```

4. Pass `tbl_review` to `show_query()` to confirm that it is pulling from a 
new temporary table

```{r}

```

5. Coerce "product" to a character

```{r}

```

6. Wrap the `as.character()` call, inside a `tolower()` call

```{r}

```

7. Add a count step, that breaks down the reviews by product

```{r}

```

## 04 - Introducing `chattr`
*Request a simple example from LLM*

1. Load the `chattr` library

```{r}

```


2. Call `chattr_app()` and then select the first model (DBRX)

```{r}

```
3. In the app type, and run:
"show me a simple example of a ggplot using the mtcars data set"

4. Copy the code into a script and test

