---
title: "Visualizations"
execute: 
  eval: true
  freeze: true
---

```{r, setup}
#| include: false

library(dplyr)
library(dbplyr)
```

## Catch up {.unnumbered}

```{r}
library(dplyr)
library(dbplyr)
library(DBI)

con <- dbConnect(
  odbc::databricks(),
  HTTPPath = "/sql/1.0/warehouses/300bd24ba12adf8e"
)

orders <- tbl(con, I("samples.tpch.orders"))
customers <- tbl(con, I("samples.tpch.customer"))
nation <- tbl(con, I("samples.tpch.nation"))

prep_orders <- orders %>% 
  left_join(customers, by = c("o_custkey" = "c_custkey")) %>% 
  left_join(nation, by = c("c_nationkey" = "n_nationkey")) %>% 
  mutate(
    order_year = year(o_orderdate), 
    order_month = month(o_orderdate)
  ) %>% 
  rename(customer = o_custkey) %>% 
  select(-ends_with("comment"),  -ends_with("key"))
```

## Auto-collect
**See how `ggplot2` auto-collects data before plotting**

1. Load `ggplot2`

```{r}
library(ggplot2)
```

2. Plot the `n_name` over `n_region_key` from the `nation` table. Use the
column geom.

```{r}
nation %>% 
  ggplot() +
  geom_col(aes(n_name, n_regionkey))
```

## Plot data

1. Using `prep_order`, pull the total sales by year (`o_totalprice`)

```{r}
prep_orders %>% 
  group_by(order_year) %>% 
  summarise(
    total_price = sum(o_totalprice, na.rm = TRUE)
  ) %>% 
  arrange(order_year)
```

2. Add to the code, a step to plot the data. Use a column geom

```{r}
prep_orders %>% 
  group_by(order_year) %>% 
  summarise(
    total_price = sum(o_totalprice, na.rm = TRUE)
  ) %>% 
  arrange(order_year) %>% 
  ggplot() +
  geom_col(aes(order_year, total_price)) 
```

3. Download the results to R to a variable called `sales_by_year`

```{r}
sales_by_year <- prep_orders %>% 
  group_by(order_year) %>% 
  summarise(
    total_price = sum(o_totalprice, na.rm = TRUE)
  ) %>% 
  collect()
```

4. Preview `sales_by_year`

```{r}
sales_by_year
```

5. Use `sales_by_year` to create the same plot

```{r}
sales_by_year %>% 
  ggplot() +
  geom_col(aes(order_year, total_price)) 
```

6. An example of what multiple iterations of the plot would result in

```{r}
breaks <- as.double(quantile(c(0, max(sales_by_year$total_price))))

breaks_labels <- paste(round(breaks / 1000000000, 1), "B")

sales_by_year %>% 
  ggplot() +
  geom_col(aes(order_year, total_price)) +
  scale_x_continuous(breaks = unique(sales_by_year$order_year)) +
  scale_y_continuous(breaks = breaks, labels = breaks_labels) +
  xlab("Year") +
  ylab("Total Sales") +
  labs(title = "Sales by year") +
  theme_light()
```

## Plot data by country

```{r}
country <- "FRANCE"
sales_by_year <- prep_orders %>% 
  filter(n_name == country) %>% 
  group_by(order_year) %>% 
  summarise(
    total_price = sum(o_totalprice, na.rm = TRUE)
  ) %>% 
  collect()
```

```{r}
breaks <- as.double(quantile(c(0, max(sales_by_year$total_price))))

breaks_labels <- paste(round(breaks / 1000000000, 1), "B")

sales_by_year %>% 
  ggplot() +
  geom_col(aes(order_year, total_price)) +
  scale_x_continuous(breaks = unique(sales_by_year$order_year)) +
  scale_y_continuous(breaks = breaks, labels = breaks_labels) +
  xlab("Year") +
  ylab("Total Sales") +
  labs(title = "Sales by year", subtitle = country) +
  theme_light()
```



